---
title: "Reassign anem_obs"
output: html_notebook
---
#Setup 
```{r setup}
library(tidyverse)
source("../genomics/scripts/field_helpers.R")
source("../genomics/scripts/lab_helpers.R")
leyte <- write_db("Leyte")
```


```{r anem_table}
# find  anem_ids
anem <- leyte %>% 
  tbl("anemones") %>%
  collect()
```
**The past method**
# which anemones already have an anem_obs number?
```{r}
have_obs <- anem %>% 
  filter(!is.na(anem_obs)) %>% 
  select(anem_id, anem_obs) %>% 
  mutate(anem_obs = as.integer(anem_obs)) %>% 
  distinct() %>% 
  arrange(anem_id)
```
# attach anem_obs to any new anem observations that have been seen before, can only do this by removing and reattaching anem_obs
```{r}
all_anem <- select(anem, -anem_obs)
all_anem <- left_join(all_anem, have_obs, by = "anem_id")
```

# which of these have an old_anem_id but no anem_obs?
```{r}
old_obs <- all_anem %>% 
  filter(!is.na(old_anem_id) & is.na(anem_obs))
```
# attach existing anem_obs to any old_anem_id that has already been assigned one
```{r}
old_obs <- select(old_obs, -anem_obs)
old_obs <- left_join(old_obs, have_obs, by = c("old_anem_id" = "anem_id")) %>% 
  distinct()
```
# update these rows in all_anem
```{r}
all_anem <- anti_join(all_anem, old_obs, by = "anem_table_id")
all_anem <- rbind(all_anem, old_obs)
rm(old_obs)
```
# which anemones have old_anem_id and no anem_obs (still)
```{r}
have_old <- all_anem %>% 
  filter(!is.na(old_anem_id) & is.na(anem_obs)) 

# assign an anem obs to anems with no obs and an old_anem_id ####
x <- max(all_anem$anem_obs, na.rm = T)
y <- x + 1
z <- x+nrow(have_old)
have_old <- have_old %>% 
  mutate(anem_obs = y:z)
# split out the old anem ids with their newly assigned anem obs
olds <- have_old %>% 
  select(old_anem_id, anem_obs) %>% 
  rename(anem_id = old_anem_id)
# do the same for anem_ids
anems <- have_old %>% 
  select(anem_id, anem_obs)  %>% 
  distinct()
# combine
new_obs <- rbind(olds, anems) %>% 
  distinct()
# add to have_obs
have_obs <- rbind(have_obs, new_obs) 
have_obs <-  have_obs %>% 
  mutate(anem_id = as.integer(anem_id),
    anem_obs = as.integer(anem_obs)) %>% 
      distinct()

# do any anem_ids have more than one anem_obs? No.
test <- have_obs %>% 
  group_by(anem_id) %>% 
  summarise(count = n()) %>% 
  filter(count > 1)

# attach to all_anem
all_anem <- select(all_anem, -anem_obs) 
all_anem <- left_join(all_anem, have_obs, by = "anem_id")
rm(new_obs, olds, anems, have_old)
```
# Test
```{r}
# Are there any repeat anem_table_ids?
test <- all_anem %>%
  group_by(anem_table_id) %>% 
  summarise(count = n()) %>% 
  filter(count > 1)

# do any anem_ids have more than one anem_obs? 
test <- all_anem %>% 
  filter(!is.na(anem_id)) %>% 
  select(anem_id, anem_obs) %>% 
  distinct() %>% 
  group_by(anem_id) %>% 
  summarise(count = n()) %>% 
  filter(count > 1)
```
# Write to the database
```{r}
# backup 
anem <- leyte %>% 
  tbl("anemones") %>% 
  collect()

write_rds(anem, path = paste0("~/Documents/db_backups/", Sys.Date(), "_anemones.Rdata"))

# write
# dbWriteTable(leyte, "anemones", all_anem, row.names = F, overwrite = T)
# dbDisconnect(leyte)
```

