---
title: "My Baits"
output: html_notebook
params:
  num_rxns: 48
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

**Part 1: Hybridization**  
Here, sequencing libraries are denatured, their adapters are blocked by adapter blocker, and baits are allowed to encounter and then hybridize to their targets.  

Gather these components:  
Reagents:  
* HYB reagents (Box 1 & 2): H1(orange) H2(red) H3(yellow) H4(teal)
* BLOCK reagents (Box 2): B1(green) B2(blue) B3(mustard)  
* RNase Block (Box 2) **Keep on ice** (purple)  
* Baits (Box 3) **Keep on ice** (white)  
* Sequencing libraries to be enriched  

*Thoroughly vortex HYB #1 H1 before use, and bring HYB #4 H4 to room temperature to fully dissolve its SDS before use*  

Equipment:  
* 1.7 mL (×2) and 50 mL (×1) nuclease-free tubes
* Low-bind 0.2mL tubes with individual caps (×2 per reaction)
* Pipettors & tips; multichannel pipettor for pipetting up to 20 μL recommended
* Vortex mixer
* Thermal cycler  

Program the thermal cycler:
Program the lid temperature to stay at 105°C, or at least 10°C above each step temperature, to keep evaporation to a minimum.

```{r echo=FALSE, message=FALSE}
a <- read_csv("step, temp_C, time
  1, 95, 5m
  2, 65, 5m
  3, 65, inf")

kable(a) %>% 
  kable_styling()
```

Assemble the Hybridization Mix, briefly vortex and centrifuge:
```{r echo=FALSE, message=FALSE}
b <- read_csv("component, uL_rxn
  hyb1, 9
  hyb2, 0.5
  hyb3, 3.5
  hyb4, 0.5
  rnase block, 1
  baits, 5.5") %>% 
  mutate(total = uL_rxn * params$num_rxns)

total <- b %>% 
  summarise(tot_per_rxn = sum(uL_rxn), 
    tot_today = sum(total)) %>% 
  mutate(component = "total") %>% 
  rename(uL_rxn = tot_per_rxn, 
    total = tot_today)

b <- rbind(b, total)

kable(b) %>% 
  kable_styling()
```

For each capture reaction, aliquot 18.5 μL of Hybridization Mix to a 0.2 mL tube. These are now referred to as “HYBs”

Assemble the Blockers Mix and briefly vortex:
```{r echo=FALSE, message=FALSE}
c <- read_csv("component, uL_rxn
  block1, 2.5
  block2, 0.5
  block3, 3.5") %>% 
  mutate(total = uL_rxn * params$num_rxns)

total <- c %>% 
  summarise(tot_per_rxn = sum(uL_rxn), 
    tot_today = sum(total)) %>% 
  mutate(component = "total") %>% 
  rename(uL_rxn = tot_per_rxn, 
    total = tot_today)

c <- rbind(c, total)

kable(c) %>% 
  kable_styling()
```
4. For each capture reaction, aliquot 5 μL Blockers Mix to a low-bind 0.2mL tube
5. Add 7 μL DNA library (100 – 500ng) to each Blockers Mix aliquot and homogenize by pipetting.

These are now referred to as “LIBs” total volume: 12 μL


**Part 2: Bind and Wash**

Prepare 10mM Tris-Cl, 0.05% Tween-20 solution (pH 8.0-8.5)
```{r echo=FALSE, message=FALSE}
sol <- read_csv("component, init_conc, fin_conc 
                mM Tris-Cl, 1000, 10
                % Tween-20, 100, 0.05") %>% 
  mutate(ul_fin_vol = 30*params$num_rxns*1.1, 
         ul_init_vol = fin_conc * ul_fin_vol/init_conc)
```

