---
title: "My Baits"
output: html_notebook
params:
  num_rxns: 48
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

##Part 1: Hybridization  

###1.1 - Getting Started
Here, sequencing libraries are denatured, their adapters are blocked by adapter blocker, and baits are allowed to encounter and then hybridize to their targets.  

**Gather these components:  **
Reagents:  
* HYB reagents (Box 1 & 2): H1(orange) H2(red) H3(yellow) H4(teal)
* BLOCK reagents (Box 2): B1(green) B2(blue) B3(mustard)  
* RNase Block (Box 2) **Keep on ice** (purple)  
* Baits (Box 3) **Keep on ice** (white)  
* Sequencing libraries to be enriched  

*Thoroughly vortex HYB #1 H1 before use, and bring HYB #4 H4 to room temperature to fully dissolve its SDS before use*  
Equipment:  
* 1.7 mL (×2) and 50 mL (×1) nuclease-free tubes
* Low-bind 0.2mL tubes with individual caps (×2 per reaction)
* Pipettors & tips; multichannel pipettor for pipetting up to 20 μL recommended
* Vortex mixer
* Thermal cycler  

**Program the thermal cycler:  **
Program the lid temperature to stay at 105 C, or at least 10 C above each step temperature, to keep evaporation to a minimum.

```{r echo=FALSE, message=FALSE}
a <- read_csv("step, temp_C, time
  1, 95, 5m
  2, 65, 5m
  3, 65, inf")

kable(a) %>% 
  kable_styling()
```
###1.2 - Mix Setups
**Assemble the Hybridization Mix, briefly vortex and centrifuge:  **
```{r echo=FALSE, message=FALSE}
b <- read_csv("component, uL_rxn
  hyb1, 9
  hyb2, 0.5
  hyb3, 3.5
  hyb4, 0.5
  rnase block, 1
  baits, 5.5") %>% 
  mutate(total = uL_rxn * params$num_rxns)

total <- b %>% 
  summarise(tot_per_rxn = sum(uL_rxn), 
    tot_today = sum(total)) %>% 
  mutate(component = "total") %>% 
  rename(uL_rxn = tot_per_rxn, 
    total = tot_today)

b <- rbind(b, total)

kable(b) %>% 
  kable_styling()
```

** For each capture reaction, aliquot 18.5 μL of Hybridization Mix to a 0.2 mL tube. These are now referred to as “HYBs” **  

**Assemble the Blockers Mix and briefly vortex:  **
```{r echo=FALSE, message=FALSE}
c <- read_csv("component, uL_rxn
  block1, 2.5
  block2, 0.5
  block3, 3.5") %>% 
  mutate(total = uL_rxn * params$num_rxns)

total <- c %>% 
  summarise(tot_per_rxn = sum(uL_rxn), 
    tot_today = sum(total)) %>% 
  mutate(component = "total") %>% 
  rename(uL_rxn = tot_per_rxn, 
    total = tot_today)

c <- rbind(c, total)

kable(c) %>% 
  kable_styling()
```
**For each capture reaction, aliquot 5 μL Blockers Mix to a low-bind 0.2mL tube.**  
**Add 7 μL DNA library (100 – 500ng) to each Blockers Mix aliquot and homogenize by pipetting.**  

These are now referred to as “LIBs” total volume: 12 μL

###1.3 Reaction Assembly
1. Put the LIBs in the thermal cycler, close the lid, and start the thermal program.  
*Double check that the lid temperature is programmed to stay at 105 C, or at least 10 C above each step temperature, to keep evaporation at a minimum.*  

* Place a strip of 8 tubes horizontally in a central location on the thermal cycler and place empty tubes at each corner to keep the lid flat while doing fewer than 2 strips worth of captures.  

2. Once the cycler reaches step 2 of the program (the hybridization temperature), pause the program, put the HYBs in the thermal cycler, close the lid and resume the program.  

3. After step 2 of the program is complete, **leaving all tubes in the thermocycler**, pipette 18 uL of each HYB to each LIB.  **Use a multichannel pipettor for easier execution**.  Gently homogenize by pipetting up and down 5 times.

4. Dispose of HYB tubes.  Close the lid of the thermal cycler and allow the reactions to incubate at hybridization temperature for your chosen time (e.g., 16 hours).


##Part 2: Bind and Wash  
*Start 90 minutes before intended hybridzation stop-time*  
Here, bait-target hybrids are bound to streptavidin-coated magnetic beads, and then non-hybridized and non-specifically hybridized DNA are removed with a series of wash steps.  

Follow this protocol if you have a 96-well magnetic particle collector. It can be used with any number of captures. For binding and washing steps, use PCR strips with individually-attached caps to enable vortexing and minimize cross-contamination. Work with up to six 8-well strips on a single 96-well particle collector at a time to enable opening the lids.  

###2.1 - Getting Started
Gather these components:  

Reagents:  

* HYB#4  

* Binding Buffer  

* Wash Buffer 2  

* Dynabeads MyOne Streptavidin C1 Beads (30 μL per reaction)  

* Nuclease-free molecular biology-grade (“NF”) water (~50 mL)  

* 10mM Tris-Cl, 0.05% TWEEN®-20 solution (pH 8.0-8.5)  

*Bring HYB #4 and Wash Buffer 2 to room temperature to dissolve SDS prior to use*  

To prepare 10mM Tris-Cl, 0.05% Tween-20 solution (pH 8.0-8.5)  

```{r echo=FALSE, message=FALSE}
sol <- read_csv("component, init_conc, fin_conc 
                mM Tris-Cl, 1000, 10
                % Tween-20, 100, 0.05") %>% 
  mutate(ul_fin_vol = 30*params$num_rxns*1.1, 
         ul_init_vol = fin_conc * ul_fin_vol/init_conc)
kable(sol) %>% 
  kable_styling()
```

Equipment:  

* 0.2 mL standard-profile PCR strips with individually-attached caps (one well per reaction)  

* Thermal cycler with heated lid  

* Nuclease-free 50 mL tubes (1 per 68 capture reactions)  

* Pipettors and tips for 20 μL – 500 μL volumes  

* STRONGLY RECOMMENDED: Multichannel pipettor for 20-200 μL  

* 96-well magnetic particle collector  

* Water bath set to hybridization temperature  

* Vortex mixer  

* Minicentrifuge with adapters for 0.2 mL tubes/strips  

###2.2 - Wash Buffer 2.2 Preparation
The following procedure makes enough Wash Buffer 2.2 for washing 68 capture reactions; scale up if needed. Wash Buffer 2.2 can be stored at 4 C for 1 month.  

1. Combine the following:
```{r echo=FALSE, message=FALSE}
sol <- read_csv("component, ul_rxn_vol
                HYB #4, 5.88
                pH2O, 582.35
                Wash Buffer 2, 147.05") %>% 
  mutate(ul_fin_vol = ceiling(ul_rxn_vol*params$num_rxns))
         
         
kable(sol) %>% 
  kable_styling()
```

Prepare in a 50 mL tube. Vortex thoroughly, label “Wash Buffer 2.2”  

2. Heat the Wash Buffer 2.2 to the hybridization temperature in the water bath for at least 45 minutes before use.  

###2.3 - Bead preparation
1. For each capture reaction, aliquot 30 μL Dynabeads® MyOneTM Streptavidin C1 beads to a 0.2 mL tube of a PCR strip with individually-attached caps.  

2. Pellet the beads in the magnetic particle collector (“MPC”) until the suspension is clear (~1-2 minutes). Leaving the tubes on the magnet, remove and discard the supernatant.  

3. Add 200 μL Binding Buffer to each bead aliquot. Vortex 3 seconds and centrifuge briefly. Pellet in the MPC for 2 minutes, remove and discard the supernatant.  

4. Repeat Step 3 above twice for a total of three washes.  

5. Resuspend each washed bead aliquot in 70 μL Binding Buffer.  

*TIP: With a MPC for ~1.5mL tubes, beads can be prepared in batches of up to 8 reactions-worth (240 μL) at a time in a 1.7 mL tube. When doing so, multiply the wash and resuspension volumes by the number of reactions in the batch. For example, for eight reactions-worth, wash three times with 1.6 mL and resuspend in 560 uL Binding Buffer, then aliquot 70 μL suspension to 0.2 mL wells for each binding reaction.*

###2.4 - Bead-bait binding
1. Heat the bead aliquots to your chosen hybridization temperature (e.g., 65 C) for 2 minutes in the thermal cycler. Set the lid temperature at least 10 C higher than the block.  

2. Transfer each capture reaction to the heated bead aliquots. Mix by pipetting.  

3. Incubate the libraries + beads in the thermal cycler at the hybridization temperature for 30
minutes. Agitate every 10 minutes by flicking the tubes and briefly centrifuging.  

###2.5 - Bead washing  

1. Pellet the beads in the MPC for 2 minutes and remove the supernatant.  

2. Add 180 μL heated Wash Buffer 2.2 to the beads, briefly vortex, and briefly centrifuge.  

3. Incubate for 10 minutes at the hybridization temperature in the thermal cycler; agitate every 3
minutes by briefly vortexing then centrifuging the strip. Pellet and remove the supernatant.  

4. Repeat step 3 three times for a total of four washes. After the final wash and pelleting, remove as much fluid as possible without touching the bead pellet.  

##Part 3: Library Elution & Amplification  
Here, the bead-bound enriched library is released from the RNA bait via heat denaturation, and then amplified in preparation for sequencing.  


###3.1 - Enriched Library Elution
1. Add 30 μL of 10mM Tris-Cl, 0.05% TWEEN®-20 solution (pH 8.0 – 8.5) to the washed beads and thoroughly resuspend by pipetting.  

**If you are using KAPA® HiFi polymerase for amplification, skip steps 2 and 3, and use this bead suspension directly in amplification. For other polymerases, continue to step 2.**

*2. Incubate the suspension at 95°C for 5 minutes.*  

*3. Pellet the beads in the MPC, and remove the supernatant, which contains the enriched library.*  

###3.2 - Suggested Amplification Setup

Prior to sequencing, Arbor recommends amplifying the enriched library using a polymerase with reduced GC and length bias, such as KAPA® HiFi (Kapa Biosystems; see Quail et al. 2012; doi:10.1038/nmeth.1814).

```{r echo=FALSE, message=FALSE}
sol <- read_csv("component, fin_conc, ul_per_rxn 
                pH2O, NA, 5
                2x KAPA HiFi HotStart ReadyMix, 1X, 25
                10uM F primer, 500, 2.5
                10uM R primer, 500, 2.5
                library, NA, 15") %>% 
  mutate(ul_fin_vol = ul_per_rxn * params$num_rxns*1.1)
kable(sol) %>% 
  kable_styling()
```

Run the "REAMP" protocol in the thermal cycler.

If there were beads in the reamp, place on magnet and keep supernatant.

Clean with Ampure and quantify.
