---
title: '`r paste("Digest plan for ", params$first, "-", params$last, sep = "")`'
params:
  id: digest_id
  first_sample: digest$extraction_id[1]
  last_sample: digest$extraction_id[91
  type: digest
output:
  pdf_document: default
---
This document is able to take a set of extracts and prepare them for digest by assigning digest_ids to them, mapping where they will go in the digest plate, and mapping where to find them in extraction plates. Currently this only fills one plate and puts the excess in tubes.  Adjust as needed.

```{r workspace, echo = FALSE, message=FALSE}
source("../../laboratory/scripts/lab_helpers.R")
library(knitr)
library(kableExtra)
library(ggplot2)
library(RColorBrewer)
lab <- read_db("Laboratory")
digest <- readRDS("samples_need_digest.Rdata")
```

**Get a list of extracts to be digested, **
It is important to have extraction_id, sample_id, well, plate, and quant.  
```{r extracts, echo = FALSE}
digest <- digest %>% 
  arrange(extraction_id)

extr <- lab %>%
  tbl("extraction") %>% 
  filter(extraction_id %in% digest$extraction_id) %>% 
  select(sample_id, extraction_id, well, plate, quant) %>% 
  collect()
```

Double check that these samples are high enough quality to be digested again.  Look at the following table for multiple digests of the same extract all with quantities lower than 2.5.  Add these extracts to the volumes rem_e extract list.
```{r}
dig <- lab %>%
  tbl("digest") %>% 
  filter(extraction_id %in% extr$extraction_id) %>% 
  select(digest_id, extraction_id, quant, notes) %>% 
  collect()
```

**How many plates will this fill?**

```{r num_plates, echo = FALSE}
nrow(extr)/96
```

The extractions are **not** going to go into the same wells they were in on the extraction plate.

Need to generate a new plate map

```{r}
  # make a dataframe of the list_of_ids
  ids <- extr %>% 
  select(extraction_id) %>% 
  arrange(extraction_id)
 
  # how many rows are in the table (how many samples)?
  y <- nrow(ids)

  # how many plates would these make, 94 samples plus 2 blanks per plate
  (nplates <- floor(y/94)) # extra parenthesis are to print

  # define wells

  well <- 1:63
  
  # insert the negative controls and set up the plate
  plate <- data.frame() # blank data frame to build upon
  i <- 1
    c <- 96*i-95 # well 1 on a plate
    d <- 96*i-85 # 11
    e <- 96*i-84 # 12 negative control well
    f <- 96*i-83 # 13
    g <- 96*i-36 # 60
    h <- 96*i-35 # 61 negative control well
    j <- 96*i-34 # 62
    k <- 96*i-33  # 65
    str1 <- as.data.frame(cbind(well[c:d], ids[c:d,])) # 1:11
    names(str1) <- c("well", "extraction_id")
    str2 <- as.data.frame(cbind(well[e], "XXXX")) # because the first blank is in the 12th position
    names(str2) <- c("well", "extraction_id")
    str3 <- as.data.frame(cbind(well[f:g], ids[e:l,])) #13:60 in plate, 12:59 in list
    names(str3) <- c("well", "extraction_id")
    str4 <- as.data.frame(cbind(well[h], "XXXX")) # because the 2nd blank is in the 61st position
    names(str4) <- c("well", "extraction_id")
    str5 <- as.data.frame(cbind(well[j:k], ids[g:k,]))# 62:96 in plate, 60:94 in list
    names(str5) <- c("well", "extraction_id")
    
    # and stick all of the rows together
    temp <- data.frame(rbind(str1, str2, str3, str4, str5))
    temp <- temp %>% 
      mutate(row = c("A", "B", "C", "D", "E", "F", "G", "H", "A", "B", "C", "D", "E", "F", "G", "H", "A", "B", "C", "D", "E", "F", "G", "H", "A", "B", "C", "D", "E", "F", "G", "H", "A", "B", "C", "D", "E", "F", "G", "H", "A", "B", "C", "D", "E", "F", "G", "H", "A", "B", "C", "D", "E", "F", "G", "H", "A", "B", "C", "D", "E", "F", "G", "H", "A"), 
             col = c(1,  1,  1,  1,  1,  1,  1,  1,  2,  2,  2,  2,  2,  2,  2,  2,  3,  3,  3,  3,  3,  3,  3,  3,  4,  4,  4,  4,  4,  4,  4,  4,  5,  5,  5,  5,  5,  5,  5,  5,  6,  6,  6,  6,  6,  6,  6,  6,  7,  7,  7,  7,  7,  7,  7,  7,  8,  8,  8,  8,  8,  8,  8,  8,  9)) %>% 
      select(-well)
    
    

  
  # put the samples in order of id (with negative controls inserted)
extr <- left_join(extr, temp, by = "extraction_id")
b1 <- c(NA, "XXXX", "D2", NA, NA, "D", 2)
b2 <- c(NA,"XXXX", "E8", NA, NA, "E", 8)
extr <- rbind(extr, b1, b2)
  
```




**Generate a digest_id**
```{r digest_id, echo = FALSE}
dig_max <- lab %>% 
  tbl("digest") %>% 
  select(digest_id) %>% 
  summarise(max = max(digest_id)) %>% 
  collect() %>% 
  mutate(max = as.numeric(substr(max, 2, 5)))

extr <- extr %>% 
  arrange(col, row) %>% 
  # create digest ids
  mutate(digest_id = 1:nrow(extr)) %>% 
  mutate(digest_id = digest_id + dig_max$max) %>% 
  mutate(digest_id = paste("D", digest_id, sep = "")) %>% 
  mutate(ng_in = 30*as.numeric(quant))
```

**Name digest plates**
```{r name_dig_plates, echo=FALSE}
dig_lim <- extr %>% 
  summarise(max = max(digest_id), 
            min = min(digest_id))

extr <- extr %>% 
  mutate(dig_plate = paste(dig_lim$min, dig_lim$max, sep = "-"), 
         well = paste0(row, col))

```
**Make plate maps of destinations**
*Maps are individual output files sent to the plots folder*

```{r dig_dest, echo = FALSE, message=FALSE}
  platex <- extr %>% 
  mutate(ng_in = ng_in/10) %>% 
  mutate(filter = 30) %>%
  mutate(filter = ifelse(ng_in >= 5000 & ng_in <= 10000, 15, filter),
         filter = ifelse(ng_in > 10000, 7, filter),
         filter = as.factor(filter)) %>%
  select(well, dig_plate, digest_id, filter)
  
  # display.brewer.all(3)
  # x <- 
    # debugonce(heatmap)
    heatmap(platex, platex$digest_id)
  
  # +
  # scale_fill_manual(values=brewer.pal(3, "Set2"), 
  #                      name="Volume\nExtract"
                    # ,
                    #    breaks=c("30", "15", "7"),
                    #    labels=c("30uL", "15uL", "7uL")
                    # )
  ggsave(paste("../procedural_notebooks/plots/", platex$dig_plate, ".pdf", sep = ""), device = "pdf", width = 11, height = 8.5)

}
```
**Make heatmap of extract destinations **
*Maps are individual output files sent to the plots folder*
```{r extr_dest, echo = FALSE, message=FALSE}
  platex <- extr %>% 
  mutate(ng_in = ng_in/10) %>% 
  mutate(filter = 30) %>% 
  mutate(filter = ifelse(ng_in >= 5000 & ng_in <= 10000, 15, filter), 
         filter = ifelse(ng_in > 10000, 7, filter), 
         filter = as.factor(filter)) %>% 
  select(well, dig_plate, extraction_id, filter)
  
  # display.brewer.all(3)
  x <- heatmap(platex, platex$extraction_id) +
  scale_fill_manual(values=brewer.pal(3, "Pastel2"), 
                       name="Volume\nExtract",
                       breaks=c("30", "15", "7"),
                       labels=c("30uL", "15uL", "7uL"))
  ggsave(paste("../procedural_notebooks/plots/", platex$dig_plate, "_extr_dest.pdf", sep = ""), device = "pdf", width = 11, height = 8.5)

}
```

**Make plate maps of source plates**
```{r echo = F, fig.width=10, message=F}
# define the source plates
sources <- extr %>%
  select(plate) %>%
  filter(plate != "XXXX") %>%
  distinct() %>%
  arrange(plate)

# create heat maps of the plates
for (i in 1:nrow(sources)){
  source <- lab %>%
    tbl("extraction") %>%
    filter(plate == sources$plate[i]) %>%
    collect() %>%
    mutate(filter = ifelse(extraction_id %in% extr$extraction_id, "yes", "no"))
  x <- heatmap(source, source$extraction_id)
  print(x)
}
```


```{r echo=F, message=F}
# write the data to the database
w_lab <- write_db("Laboratory")
dig <- extr %>% 
  select(-plate, -row, -col, -sample_id) %>% 
  rename(plate = dig_plate)
  
# dbWriteTable(w_lab, "digest", dig, append = T, row.names = F)
# dbDisconnect(w_lab, lab)
```



**Prepare digest plate by adding specified amount of sample to each well according to the digest maps prepared above.  If you are not continuing on to the digest step now, seal the plate with film, label it, and put it in the fridge.**

The map files can be deleted from the plots folder once they have been saved in your lab notebook.





