---
title: "Which samples still need work as of `r Sys.Date()`"
output:
  pdf_document: default
  html_notebook: default
---

Investigate the samples that need to be sequenced. Do they have any juice left?

```{r}
source("../genomics/scripts/lab_helpers.R")
load("digs_need_seq.Rdata")

# get the work history
need <- work_history(need, "sample_id")
```
```{r}

# how many times have these extracts been digested?
temp <- need %>% 
  select(sample_id, extraction_id, digest_id)
digs <- temp %>% 
  group_by(extraction_id, digest_id) %>% 
  distinct() %>%
  group_by(extraction_id) %>% 
  summarise(num_digs = n())

need <- left_join(need, digs, by = "extraction_id")

# how many times have the digests been ligated?
temp <- need %>% 
  select(sample_id, extraction_id, digest_id, ligation_id)
ligs <- temp %>% 
  group_by(digest_id) %>% 
  summarise(num_ligs = n())

need <- left_join(need, ligs, by = "digest_id")

# get the volume of digest used in the ligation
lab <- read_db("Laboratory")
ligs <- lab %>%
  tbl("ligation") %>% 
  filter(ligation_id %in% need$ligation_id) %>% 
  select(ligation_id, vol_in) %>% 
  collect()

need <- left_join(need, ligs, by = "ligation_id")

# some samples should've been removed because they say empty in the notes but apparently werent.

extr <- lab %>%
  tbl("extraction") %>% 
  filter(extraction_id %in% need$extraction_id) %>% 
  select(extraction_id, notes) %>% 
  collect()

empty_e <- extr %>% 
  filter(grepl("empty", notes))

need <- anti_join(need, empty_e, by = "extraction_id")
need <- left_join(need, extr, by = "extraction_id") %>% 
  rename(extr_notes = notes)

# get digest notes
dig <- lab %>% 
  tbl("digest") %>% 
  filter(digest_id %in% need$digest_id) %>% 
  select(digest_id, notes) %>% 
  collect()
need <- left_join(need, dig, by = "digest_id") %>% 
  rename(dig_notes = notes)

need <- need %>% 
  filter(!grepl("Bad digest", dig_notes))

```
Some digests have been ligated too many times to have enough volume left to try to ligate again:
```{r}
rem_e <- c("E0266", "E1699", "E0501", "E0525", "E0526", "E0536", "E0540", "E0569", "E0683", "E0686", "E0687", "E0703", "E0713", "E2292", "E0554", "E0741", "E0745", "E0763", "E0795", "E0797", "E0851", "E0885", "E0888", "E0891", "E0964", "E0965", "E0979", "E0980", "E0981", "E1011", "E1036", "E1338", "E1339", "E1347", "E1370", "E2759", "E2775", "E2799", "E2773", "E2763", "E2569", "E2642", "E2738", "E2707", "E2736", "E2753", "E2734", "E2735", "E2737", "E0404", "E0533", "E0555", "E0572", "E0677", "E0725", "E0793", "E1304", "E1306", "E1309", "E1311", "E1313", "E1315", "E1316", "E1319", "E1320", "E1321", "E1322", "E1323", "E1326", "E1327", "E1328", "E1329", "E1332", "E1333", "E1336", "E2690", "E2737") # these samples were digested multiple times and never had enough DNA to sequence successfully
rem_d <- c("D2511", "D2514", "D2520", "D2527", "D1836", "D2557", "D2622", "D2673", "D2674", "D2678", "D0673", "D1901", "D0697", "D0729", "D0735", "D1056", "D1956", "D1260", "D1961", "D1263", "D1963", "D1266", "D1966", "D1267", "D1967", "D2236", "D1206", "D2256", "D1207", "D2257", "D1115", "D2258", "D1118", "D2259", "D1119", "D2260", "D1212", "D2261", "D1120", "D2262", "D1217", "D2263", "D1220", "D2894", "D2999", "D2890", "D3214", "D3223", "D2793", "D2891", "D3130", "D2683", "D2906", "D2816", "D2688", "D2817", "D2736", "D2696", "D2922")  # these samples were digested more than once and the other digest can be ligated.
rem_l <- c("L0214", "L1161", "L0216", "L0223", "L0227", "L0228", "L0229", "L0233", "L0234", "L1170", "L0235", "L0237", "L0241", "L0248", "L0251", "L0252", "L0253", "L0834", "L0310", "L0332", "L1306", "L0361", "L0367", "L0836", "L0838", "L0839", "L0378", "L0381", "L0840", "L0396", "L0399", "L0648", "L0404", "L1307", "L0842", "L0843", "L0850", "L0828", "L0638", "L0830", "L0968", "L1012", "L0162", "L1086", "L0848", "L0180", "L0766", "L0746", "L0548", "L1287", "L0938", "L1011", "L1327", "L1298", "L1328", "L0767", "L0970", "L0971", "L0940", "L0972", "L0941", "L1334", "L2868", "L2630") # these samples were ligated more than once and need to be digested again.  The other ligation is kept as a flag to redigest.

need <- need %>% 
  filter(!digest_id %in% rem_d, 
         !extraction_id %in% rem_e, 
         !ligation_id %in% rem_l) %>% 
  filter(!grepl("not enough", dig_notes))
```
Basic decision making:

If an extract was digested several times and they all failed, the extract was removed.

If a digest was ligated several times and they all failed, one line was kept as a flag and that extract was marked for re-digest.


Looking through the need list, if a sample was represented by several ligations, they usually only had one digest. Those samples I removed all but the highest number ligation and set those aside to digest again.  Others didn't have a ligation and had a very high digest number.  Those were the ones that we missed by the robot this round.  I removed all but the highest digest number and set those aside to be loaded into ligation plates.  The numbers above were the multiple lines from samples that were already represented in one of the cases above.



APCL13_378, E0501 has been digested twice and both times has been too low to ligate.

APCL12_045, E1710, D4721 - set up for ligation

APCL13_123, D0461 was ligated 3 times and sequenced twice and still did not produce good sequenced.  The extract it comes from, E0248, was digested 2 times.  Could try to digest again.

APCL13_141, E0266, D0479, L0832, P024 sequenced poorly.  It was redigested as D5309 and there was not enough DNA to ligate at 50ng.  D5309 was a low concentration digest plate, this did not work to produce product.

APCL13_142, E0267, D0480, L0180, L1304, L0848 sequenced poorly every time and there isn't enough volume of extract left to digest again - double check this.

APCL13_177, E0301, D0514, L1160, P030 and L0214 in pool 11 sequenced poorly.  Try to digest again.

APCL13_179, E0303 - digest again

APCL13_186, E0310 - digest again

APCL13_190, E0314 - digest again

APCL13_191, E0315 - digest again

APCL12_034, E1699, D5220 - not enough DNA to ligate at 50ng


Need to make a list of samples that have been determined to be too low quality to sequence and make sure to note if they have tags, those fish need another fin clip.

#Separate out the list into samples that need to be digested and samples that can be set up for ligation now.
```{r}
digest <- need %>% 
  filter(!is.na(ligation_id) | digest_id < "D3479")

saveRDS(digest, file = "samples_need_digest.Rdata")

ligate <- need %>% 
  filter(!sample_id %in% digest$sample_id)
saveRDS(ligate, file = "samples_ready_ligate.Rdata")
```

