---
title: "Extraction plan for `r params$first` - `r params$last`"
output:
  html_document:
    df_print: paged
  pdf_document: default
date: "`r Sys.Date()`"
params:
  first: APCL18_659
  last: APCL18_732
  num_samples: 76
---
```{r workspace, echo=FALSE, message=FALSE}
library(knitr)
library(kableExtra)
source("../genomics/scripts/lab_helpers.R")
leyte <- read_db("Leyte")
lab <- read_db("Laboratory")
```

Organize samples into plate format: 

```{r data, echo=FALSE}
# this code gets all sample_ids within the range stated in the params above.

# connect to clownfish table in the db
work <- leyte %>% tbl("clownfish") %>%
  # choose only the sample_id column
  select(sample_id) %>% 
  # choose only the rows that meet our requirements
  filter(sample_id >= params$first, 
    sample_id <= params$last, 
    !is.na(sample_id)) %>% 
  # make sure there aren't any duplicates
  distinct(sample_id) %>% 
  # pull the data into R
  collect()

# define wells
plate <- data.frame(row = rep(LETTERS[1:8], 12), col = unlist(lapply(1:12, rep, 8))) %>% 
  mutate(sample_id = ifelse(row == "D" & col == 2, "XXXX", NA), 
    sample_id = ifelse(row == "E" & col == 8, "XXXX", sample_id))

# if the plate is not full
x <- nrow(work) + 2
if (x < 96) {
  plate <- plate %>% 
    slice(1:x)
}

# move all of the non-blank wells into their own table
samples <- plate %>% 
  filter(is.na(sample_id)) %>% 
  select(-sample_id)

# remove the non-blank wells from the plate table
plate <- anti_join(plate, samples, by = c("row", "col"))

# join the work sample_ids to the sample table well definitions
samples <- cbind(samples, work)

# rejoin the samples into the plate with the blanks, resulting in a plate of defined wells for samples and blanks.
plate <- rbind(plate, samples) %>% 
  arrange(col, row)

```
1. Sample_id plate locations
```{r sample_map, echo=FALSE, message=FALSE}
platemap <- as.matrix(reshape2::acast(plate, plate$row ~ plate$col), value.var = plate$sample_id)
knitr::kable(platemap, booktabs = T) %>% 
  # use scale_down to get map to fit within the bounds of the pdf
  kable_styling(latex_options = "scale_down")
```
```{r extr_ids, echo=FALSE}
### ONLY DO THIS ONCE ###
extracted <- lab %>% tbl("extraction") %>% 
  summarise(last = max(extraction_id, na.rm = T)) %>% 
  collect() %>% 
  mutate(last = substr(last, 2,5))

plate <- plate %>% 
  mutate(well = 1:nrow(plate)) %>% 
  mutate(extraction_id = paste("E", well + as.numeric(extracted$last), sep = "")) %>% 
  mutate(well = paste(row, col, sep = "")) %>% 
  mutate(method = "Ali", 
    final_vol = "45")

plate_name <- plate %>% 
  summarise(first = min(extraction_id), 
    last = max(extraction_id, na.rm = T))

plate <- plate %>% 
  mutate(plate = paste(plate_name$first, plate_name$last, sep = "-"))
```
2. Extraction_id plate locations
```{r extr_map, echo=FALSE, message=FALSE}
# created map table of only 3 columns becaues acast is having trouble parsing out which columns to use
map <- plate %>%
  select(row, col, extraction_id)
platemap <- as.matrix(reshape2::acast(map, map$row ~ map$col), value.var = map$extraction_id)

# remove row and col from plate
plate <- select(plate, -row, -col)

knitr::kable(platemap, booktabs = T) %>% 
  kable_styling()

```
Write the newly assigned extraction_ids to the database
```{r write}
lab <- write_db("Laboratory")
# dbWriteTable(lab, "extraction", plate, row.names = F, overwrite = F, append = T)
dbDisconnect(lab)
rm(lab)
```

**Prepare a solution of Lifton's Buffer**
```{r liftons, echo=FALSE}
liftons <- readr::read_csv("ingredient, initial_conc, initial_vol_mL, final_conc, final_vol_mL
mM_EDTA, 500, NA, 100, NA
mM_Tris-HCl_pH7.5, 1000, NA, 25, NA
perc_SDS, 20, NA, 1, NA
pH2O, NA, NA, NA, NA
") %>% 
  mutate(final_vol_mL = 250, 
         initial_vol_mL = final_conc* final_vol_mL/initial_conc,
         initial_vol_mL = ifelse(ingredient == "pH2O",final_vol_mL - sum(initial_vol_mL, na.rm = T), initial_vol_mL))
 
         
kable(liftons) %>%
  kable_styling() %>%
  add_header_above(c("Lifton's Buffer" = 5))
```

**Prepare Lifton's buffer with proteinase K**
```{r enzyme, echo=FALSE}
enzyme <- readr::read_csv("ingredient, initial_conc, initial_vol_uL, final_conc, final_vol_uL
mM_DTT, 1000, NA, 75, NA
mg/ml_ProteinaseK, 20, NA, 4.2, NA
Liftons, NA, NA, NA, NA
") %>% 
  mutate(final_vol_uL = 3500, 
         initial_vol_uL = final_conc*as.numeric(final_vol_uL)/initial_conc, 
         initial_vol_uL = ifelse(ingredient == "Liftons",as.numeric(final_vol_uL) - sum(initial_vol_uL, na.rm = T), initial_vol_uL))

kable(enzyme) %>%
  kable_styling() %>%
  add_header_above(c("Lifton's with Proteinase K" = 5))
```
**Prepare hybridization solution**
```{r hybrid, echo=FALSE}
hybrid <- readr::read_csv("ingredient, initial_conc, initial_vol, final_conc, final_vol
M_NaCl, 4.45, NA, 2.5, NA
perc_PEG_8000, 50, NA, 20, NA
mM_DTT, 1000, NA, 25, NA
pH2O, NA, NA, NA, NA
") %>% 

  mutate(final_vol = 15000,
         initial_vol = final_conc*final_vol/initial_conc, 
         initial_vol = ifelse(ingredient == "pH2O",final_vol - sum(initial_vol, na.rm = T), initial_vol))

kable(hybrid) %>%
  kable_styling() %>%
  add_header_above(c("Hybridization Buffer" = 5))
```
**Prepare low TE solution**
```{r low_TE, echo=FALSE}
low_te <- readr::read_csv("ingredient, initial_conc, initial_vol, final_conc, final_vol
mM_Tris-HCl_pH7.5, 1000, NA, 10, NA
mM_EDTA, 500, NA, 0.1, NA
pH2O, NA, NA, NA, NA") %>% 
  mutate(final_vol = 15000, 
         initial_vol= final_conc*final_vol/initial_conc, 
         initial_vol = ifelse(ingredient == "pH2O",final_vol - sum(initial_vol, na.rm = T), initial_vol))

kable(low_te) %>%
  kable_styling() %>%
  add_header_above(c("Low TE buffer" = 5))
```
**Load samples for lysis**  

1. Add 80uL Lifton's buffer to each well of a 96 well plate.  
2. Add fin clips measuring 2-25mm^2 to each well.  
3. Add 40uL enzyme solution to each well, pipet up and down to mix.  
4. Seal plate and incubate at 55C overnight.  
  
  
**Extract DNA from lysate**  
1. To a new plate add 45uL hybridization buffer to each well.  
2. Add 15uL Agencourt Ampure beads.  
3. Add 45uL of the lysate and pipet up and down to mix.  
4. Incubate at room temperature for 5 minutes, then place on magnet and allow the wells to clear.  
5.  Remove supernatant.  
6. Remove from the magnet and use 150uL freshly prepared 80% EtOH to resuspend the beads.  
7. Wash with ethanol again  
8. wash with ethanol a 3rd time.  
9. Allow beads to air dry on the magnet   
10. Add 45uL low TE to elute DNA from the beads  

