---
title: Working with adapters for `r params$ng_DNA`ng pools created with PstI-MluCI on
  `r Sys.Date()`
params:
  ng_DNA: 50
  vol_P1: 25
  vol_P2: 500
  rehydrate: FALSE
  anneal: FALSE
  working: TRUE
  
output:
  pdf_document: default
---
This document walks through making working adapters for ligation.



**Enter values into the parameters above:  
ng_DNA: what is the input ng for the intended pool?  
vol_P1: how much P1 adapter do you want to make (25 is default)  
vol_P2: how much P2 adapter do you want to make (refer to your ligate script to see how much will be needed, ~212 for 96 well plate)  
rehydrate: do you need to rehydrate oligos?  
anneal: do you need to anneal adapters?  
working: do you need to create working stock?**  

```{r echo=FALSE, message=FALSE}
library(dplyr)
```


```{r rehydrate, echo=FALSE, message=FALSE}
library(dplyr)
library(knitr)
library(kableExtra)
print(params)

if (params$rehydrate == TRUE){
   # P2.1 from 2016-04-26 has 51.4nmoles
x <- 51.4 * 5
   cat(paste("Add ", x, "uL of 1xTE buffer to the X.1 IDT tube.", sep = ""))
   # P2.2 from 2016-04-26 has 78.6nmoles
(x <- 78.6 * 5)
   cat(paste("Add ", x, "uL of 1xTE buffer to the X.2 IDT tube.", sep = ""))
}
```
```{r anneal, echo=FALSE, message=FALSE}
if (params$anneal == TRUE){
  anneal <- readr::read_csv("ingredient, vol
X.1_adapter, NA
X.2_adapter, NA
10x_buffer, NA
pH2O, NA
") %>% 
  # multiply number of samples by 1.1 for pipetting error
  mutate(
    vol = ifelse(ingredient == "X.1_adapter", 40, vol),
    vol = ifelse(ingredient == "X.2_adapter", 40, vol), 
    vol = ifelse(ingredient == "10x_buffer", 20, vol), 
    vol = ifelse(ingredient == "pH2O", 100, vol)
    )
 
kable(anneal) %>%
  kable_styling(latex_options = c("striped"), font_size = 14) %>%
  add_header_above(c("Annealed stock" = 2))
}

```

```{r calc, echo=FALSE, message=FALSE}
ng <- params$ng_DNA * .001 # convert ng to ug

# calculate ends/sample in pmoles (418544 is from bioanalyzer data using the ligation molarity calculator, APCL-PstI specific)
B10 <- 1000000000000*2*((ng/1000000)/418544.5211)
B15 <- 10
B16 <- B10 * B15
B17 <- 2
B18 <- B16/B17
B19 <- 40
B20 <- B19/B18

# calculate ends/sample in pmoles (137,165.52 is from bioanalyzer data using the ligation molarity calculator, APCL-MluCI specific)
C10 <- 1000000000000*2*((ng/1000000)/137165.52)
C15 <- 10
C16 <- C10 * C15
C17 <- 2
C18 <- C16/C17
C19 <- 40
C20 <- C19/C18

```
To make annealed adapters, combine the above ingredients and place in the thermocycler with the following temperature regimen:
incubate @ 97.5C 2.5 min; cool @ 3 per minute to 21C, hold @ 4C 


To make P1 working stock:
```{r P1mix, echo=FALSE}
mix <- readr::read_csv("ingredient, vol
P1_adapter_stock, NA
1x_annealing_buffer, NA
") %>% 
  # multiply number of samples by 1.1 for pipetting error
  mutate(
    vol = ifelse(ingredient == "P1_adapter_stock", round(params$vol_P1/B20) , vol))
mix <- mix %>% 
  mutate(
    vol = ifelse(ingredient == "1x_annealing_buffer", params$vol_P1 - as.numeric(mix$vol[1]), vol)
    )
 
kable(mix) %>%
  kable_styling(latex_options = c("striped"), font_size = 14) %>%
  add_header_above(c("P1 200ng working stock" = 2))
```
1. Prepare 2 centrifuge tubes of 1x annealing buffer  
2. Pipet 150uL of 1x annealing buffer into each well of an 8 well strip.  
3. Pipet the above amount of 1x annealing buffer into the first 48 wells of a  96 well plate.  
4. Add the above amount of each annealed P1 adapter into the corresponding wells of the plate.  
5. Vortex and spin down.  

To make P2 working stock: 
```{r echo=FALSE, message=FALSE}

P2mix <- readr::read_csv("ingredient, vol
P2_adapter_stock, NA
buffer, NA
") 
P2mix <- P2mix %>% 
  # multiply number of samples by 1.1 for pipetting error
   mutate(vol = ifelse(ingredient == "P2_adapter_stock", round(params$vol_P2/C20) , vol)) 

P2mix <- P2mix %>% 
  mutate(vol = ifelse(ingredient == "buffer", params$vol_P2 - as.numeric(P2mix$vol[1]), vol))

kable(P2mix) %>%
  kable_styling(latex_options = c("striped"), font_size = 14) %>%
  add_header_above(c("P2 200ng working stock" = 2))

```
1. Add the above amount of  1x annealing buffer to a centrifuge tube.  
2. Add the above amount of annealed P2 adapter.  
3. Vortex and spin down.





