---
title: "PCR pools on `r Sys.Date()`"
params:
  num_negs: 0
  num_wells: 96
  rxn_size_ul: 20
output:
  pdf_document: default
  html_notebook: default
---
1. Update db with PCR numbers and indices.  
```{r update_db, echo=FALSE, message=FALSE}
source("../genomics/scripts/lab_helpers.R")
library(tidyverse)
# lab <- write_db("Laboratory")
P <- paste0("P", formatC(73:128, width = 3, format = "d", flag=0))
pools <- tibble(pcr_id = P) %>% 
  filter(pcr_id < "P079" | pcr_id > "P082") %>% 
  mutate(date = "2018-08-22", 
    buffer = 30, 
    size = 375, 
    vol_from_pippin = 40, 
    pool_id = paste0("Pool", substr(pcr_id, 2, 5)))

# send data to pool table in db
for_db <- pools %>% 
  select(-pcr_id)

# dbWriteTable(lab, "pool", for_db, row.names=F, append=T)
rm(for_db)

# send data to pcr table in db
pools <- pools %>% 
  select(-buffer, -vol_from_pippin, -size) %>%
  # date of pcr
  mutate(date = "2018-09-04", 
    # 4 pcr reactions for each pool
    rxns = 4, 
    # add illumina index number
    index = rep(1:12, length = nrow(pools)), 
    # 20uL pcr reactions
    vol_rxns = params$rxn_size_ul, 
    # 430 bp for APCL project
    bp = 430)

# dbWriteTable(lab, "pcr", pools, row.names=F, append = T)

```



PCR pools
```{r pcr setup}
library(dplyr)
num_wells <- params$num_wells

neg_control <- params$num_negs

rxns_plus_error <- ceiling((num_wells + neg_control)*1.1)

rxn_size <- params$rxn_size_ul

final_volume <- rxn_size * rxns_plus_error

calc_table <- readr::read_csv("item, vol_to_add, initial_conc, final_conc, final_vol, per_rxn
                              5x buffer, NA, 5, 1, NA, 4
                              F primer, NA, 10, 0.5, NA, 1
                              dNTPs, NA, 10, 0.2, NA, 0.4
                              MgCl2, NA, 50, 1.25, NA, 0.5
                              phusion, NA, NA, NA, NA, 0.2 
                              pH2O, NA, NA, NA, NA, 2.9") %>% 
  mutate(final_vol = final_volume, 
    vol_to_add = (final_conc * final_vol)/initial_conc, 
    vol_to_add = ifelse(item == "phusion", final_volume/100, vol_to_add))

added <- calc_table %>% 
  summarise(s = sum(vol_to_add, na.rm = T))

calc_table <- calc_table %>% 
  mutate(vol_to_add = ifelse(item == "pH2O", final_vol - (added$s+rxns_plus_error+(rxns_plus_error*rxn_size*.5)), vol_to_add))

# double check the match
check <- calc_table %>% 
  mutate(check = per_rxn*rxns_plus_error)

# These should all be true
floor(check$check) == floor(check$vol_to_add)

knitr::kable(calc_table) %>% 
  kableExtra::kable_styling()
```


2. Run each pool in 4 rxns at 10µL of sample each in 20µL reactions.  

3. Don’t include R primer in master mix, add later…example recipe looks like:  
* Place 9µL master mix in each well  
* Place 10µL sample in each well  
* Place 1µL R primer (of the correct index) in each well  

4. Run on thermocycler PHUSION_PCR which is 98 degrees for 30 sec, (98degrees for 10 sec, 64degrees for 30 sec, 72degrees for 30 sec) X 12, 72degrees for 10 minutes, hold at 4degreesC


After the thermocycler finishes (~35 minutes) Combine the reactions of each pool individually in tubes (o-ring if not cleaning immediately) labeled on the top with the pool number and on the side with
“PCR #" not clean date

Clean the PCR products separately (separate pools) each pool should be 80µL add 120µL Ampure - elute into 40µL TE

Qubit the pools  
Make a note of the supplies used and update the inventory

For PCR Single reaction volumes: 2.9 pH2O 4.0 Phusion buffer 0.5 MgCl2 0.4 10mM dNTPs 1.0 F Primer 1.0 R Primer (2_2) 0.2 Phusion polymerase








```{r make working stock}

```

To make working stock primers from primary stock:
For 1 plate of PCR rxns, make 100µL:
* final: 100µL @ 10µM = initial 5µL @ 200µM
* Combine 95µL pH2O + 5µL primary primer stock (in primer freezer box in freezer)
3/4/2014 - pg 54 of lab notebook
