---
title: "Rebuilding samples based on robot files"
output: html_notebook
---

Because the robot documents were re-generated in Princeton and may have changed from the database version, I am re-creating the rows for the database based on the most recent version of the robot files.

**Import the data from the robot files**
```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(knitr)
source("../genomics/scripts/lab_helpers.R")
files <- list.files("/Volumes/USB DISK/robot files/")

rows <- tibble()
for (i in files){
  dat <- read_csv(paste0("/Volumes/USB DISK/robot files/", i)) %>% 
    mutate(file = i)
  rows <- rbind(rows, dat)
}
rows <- distinct(rows) %>% 
  select(-X1)
rm(x, dat)
```
Double check that all of the pool names look correct and that there are 48 samples in each pool
```{r echo=FALSE, message=FALSE}
pools <- rows %>% 
  group_by(pool) %>% 
  summarise(count = n())
```
Pools P124 and P102 have too many samples, what is going on?
```{r echo=FALSE, message=FALSE}
strange <- rows %>% 
  filter(pool == "P124" | pool == "P102")
kable(strange)
```
P124 has duplicate rows because the file was adjusted into a new file - I just needed to run those last few rows again, I think because of a tip issue (had to re-arrange tips in the box)
```{r echo=FALSE, message=FALSE}
weird <- strange %>% 
  filter(pool == "P102") %>% 
  select(digest_id) %>% 
  distinct()
kable(weird)
```
It appears there are some duplicate digest_ids within P102 that are making it appear like more than 48 samples when really it is only 48.  However, it is possible that these samples were filled twice...D4496-D4431
```{r echo=FALSE, message=FALSE}
weirder <- strange %>% 
  filter(pool == "P102") %>% 
  group_by(digest_id) %>% 
  summarise(num = n())

weirder <- strange %>% 
  filter(digest_id >= "D4396" & digest_id <= "D4431", 
    file != "2018-08-039_a.csv")
```
Remove the duplicate rows
```{r echo=FALSE, message=FALSE}
rows <- rows %>% 
  filter(file != "2018-08-0320_c_orig.csv")

rows <- anti_join(rows, weirder)
```


Remove the columns that will not go into the database
```{r echo=FALSE, message=FALSE}
rows <- rows %>% 
  select(digest_id, dest_well, uL_in, water, DNA, pool) %>% 
  rename(vol_in = uL_in, 
    well = dest_well)

rows <- distinct(rows)

rows <- rows %>%
  mutate(num = as.numeric(substr(pool, 2,4)), 
    num = formatC(as.numeric(num), width = 3, format = "d", flag = 0), 
    pool = paste0("P", num))
rows <- rows %>% 
  mutate(date = "2018-08-21") %>% 
  select(-num)

# add ligation_id
wlab <- write_db("Laboratory")
lig <- dbReadTable(wlab, "ligation") %>% 
  select(ligation_id) %>% 
  summarize(max = max(ligation_id)) 

x <- as.numeric(substr(lig$max, 2, 5))
possibilities <- tibble(id = 1:9999) %>% 
  mutate(id = formatC(id, width = 4, format = "d", flag = 0), 
    id = paste0("L", id))

used <- dbReadTable(wlab, "ligation") %>% 
  select(ligation_id)

possibilities <- possibilities %>% 
  filter(!id %in% used$ligation_id)

# attach available ligation ids to current work
possibilities <- possibilities %>% 
  slice(1:nrow(rows))
rows <- cbind(rows, possibilities) 

rows <- rows %>% 
  rename(ligation_id = id)
```
Write the data to the database
```{r echo=FALSE, message=FALSE}


# dbWriteTable(wlab, "ligation", rows, row.names = F, append = T)
# dbDisconnect(wlab)
```

