---
title: "Check GPSSurveys2017.xls for type-os"
author: "Michelle Stuart"
date: "6/7/2017"
output:
  html_document:
    keep_md: yes
  html_notebook: default
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE)
library(tidyverse)
library(stringr)
excel_file <- ("GPSSurveys2017.xlsx")
problem <- data.frame()
```
Import the diveinfo sheet to be checked
```{r}
surv <- readxl::read_excel(excel_file, sheet = "diveinfo", col_names=TRUE)
names(surv) <- stringr::str_to_lower(names(surv))
```

Make sure that the sites on the diveinfo sheet are spelled correctly.  First, define the list of sites. Then find all of the sites that were spelled correctly in a table called good. Then find all of the sites that are in the excel file, but are not in the good table (therefore not spelled correctly).  Get rid of any that are blank lines. Add a note to any type-os found in this section and add them to the final problem list.

```{r, message=FALSE}
sites <- c("Palanas", "Wangag", "Magbangon", "Cabatoan", "Caridad Cemetery", "Caridad Proper", "Hicgop South", "Sitio Tugas", "Elementary School", "Sitio Lonas", "San Agustin", "Poroc San Flower", "Poroc Rose", "Visca", "Gabas", "Tamakin Dacot", "Haina", "Sitio Baybayon")
good <- dplyr::filter(surv, site %in% sites)
bad <- dplyr::anti_join(surv, good)
bad <- dplyr::filter(bad, !is.na(divenum))
if (nrow(bad) > 0){
  bad$typeo <- "fix site spelling on diveinfo table"
}
(problem <- rbind(problem, bad))
rm(good, bad)
```
Import the anemone sheet to be checked, remove any blank lines and any duplicated lines

```{r, warning=FALSE}
anem <- readxl::read_excel(excel_file, sheet = "anemones", col_names = TRUE, na = "")
names(anem) <- str_to_lower(names(anem))
anem <- filter(anem, !is.na(id))
anem <- distinct(anem)
```

Look for any anemspp field that is not on our defined list, make a note and add it to the problem list.
```{r, message=FALSE}
anems <- c("ENQD", "STME", "HECR", "HEMG", "STHD", "HEAR", "MADO", "HEMA", "STGI", "????", "EMPT")
good <- filter(anem, anemspp %in% anems)
bad <- anti_join(anem, good)
bad <- filter(bad, !is.na(anemspp))
if (nrow(bad) > 0){
  bad$typeo <- "fix anem spp on anem table"
}
(problem <- rbind(problem, bad))
rm(good, bad)
```
Look for any spp field that is not on our defined list, make a note and add it to the problem list.
```{r, message=FALSE}
fish <- c("APCL", "APOC", "APPE", "APSE", "APFR", "APPO", "APTH", "PRBI", "NA")
good <- filter(anem, spp %in% fish)
bad <- anti_join(anem, good) # wait for the next line before checking bad
bad <- filter(bad, !is.na(spp))
if (nrow(bad) > 0){
  bad$typeo <- "fix fish spp on anem table"
}
(problem <- rbind(problem, bad))
rm(good, bad)
```
Import the clownfish sheet to be checked, remove any blank lines and any duplicated lines

```{r}
clown <- readxl::read_excel(excel_file, sheet = "clownfish", col_names = TRUE, na = "")   
names(clown) <- stringr::str_to_lower(names(clown))
clown <- filter(clown, !is.na(divenum))
```

Look for any spp field that is not on our defined list, make a note and add it to the problem list.
```{r, message=FALSE}
good <- filter(clown, spp %in% fish)
bad <- anti_join(clown, good)
bad <- filter(bad, !is.na(spp))
if (nrow(bad) > 0){
  bad$typeo <- "fix fish spp on fish table"
}
(problem <- rbind(problem, bad))
```
Look for any col field that is not on our defined list, make a note and add it to the problem list.
```{r, message=FALSE}
colors <- c("YE", "O", "YR", "YP", "Y", "W", "WR", "WP", "BW", "B")
good <- filter(clown, color %in% colors)
bad <- anti_join(clown, good)
bad <- filter(bad, !is.na(color))
if (nrow(bad) > 0){
  bad$typeo <- "fix tail color on fish table"
}
(problem <- rbind(problem, bad))
```

Are there any anems on the clownfish sheet that aren't on the anem survey?

```{r, message=FALSE}
dive <- clown$divenum
dive <- unique(dive)

for (i in 1:length(dive)){
  fishanem <- filter(clown, divenum == dive[i]) # for one dive at a time in the fish table
  anemanem <- filter(anem, divenum == dive[i]) # and the anem table
  good <- filter(fishanem, anemid %in% anemanem$anemid) # find all of the anems that are in the anem table
  bad <- anti_join(fishanem, good) # and those anems that aren't
  bad <- filter(bad, anemid != "????") # except for any with an unknown anem 
  if (nrow(bad) > 0){
    bad$typeo <- "anemid in fish table doesn't match anem data table"
  }
  problem <- rbind(problem, bad)
}
problem
```

Import tags from scanner
```{r}
infile <- "BioTerm.txt" 
pit <- read_csv(infile, 
  col_names = c("city", "tagid", "date", "time"), 
  col_types = cols(
    city = col_character(),
    tagid = col_character(),
    date = col_character(), # have to specify as string  
    time = col_character() # have to specify as string
  )
)
pit <- distinct(pit)
pit <- as_tibble(pit) %>% # merge fields into tag number for fish
  unite(scan, city, tagid, sep = "")

pit <- filter(pit, substr(date, 7,8) == "17" & substr(date, 1,2) != "16") # find only this year

# get rid of test tags
pit <- filter(pit, substr(scan,1,3) != "989" & substr(scan,1,3) != "999")
pit <- arrange(pit, date, time)
```

Change pit tags on clownfish data sheet to reflect the whole number and select only the tagid column
```{r}
clown <- clown %>% 
  mutate(tagid = stringr::str_replace(tagid, "985_", "985153000")) %>% 
  mutate(tagid = stringr::str_replace(tagid, "986_", "986112100")) %>% 
  mutate(tagid = stringr::str_replace(tagid, "982_", "982000411"))

tagids <- clown %>% select(contains("tag")) %>% filter(!is.na(tagid))
tagids <- distinct(tagids)
```

What tags are in excel that were not scanned by the scanner (type-os) - should return 0 rows
```{r}
anti_join(tagids, pit, by = c("tagid" = "scan"))
```

What tags are in the scanner that are not in excel (type-os) - should return 0 rows
```{r}
anti_join(pit, tagids, by = c("scan" = "tagid"))  
```
```{r}
problem
```




