---
title: "Trim GPX for QGIS"
author: "Michelle Stuart"
date: "6/7/2017"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse) # for pipe functions
library(lubridate) # for changing time zones
library(stringr) # for combining dates and times into dttm
source("readGPXGarmin.R")
source("writeGPXGarmin.R")
```
  
Get the dive info from the excel spreadsheet
```{r}
excel_file <- ("GPSSurveys2017.xlsx")
surv <- readxl::read_excel(excel_file, sheet = "diveinfo", col_names=TRUE)
names(surv) <- stringr::str_to_lower(names(surv))
```
  
Reformat to reduce the number of columns and get rid of weird dates from excel
```{r}
surv <- surv %>% 
  select(divenum, date, starttime, endtime, pausestart, pauseend, gps)

surv <- surv %>% 
  separate(starttime, into = c("baddate", "starttime"), sep = " ") %>%
  separate(endtime, into = c("baddate2", "endtime"), sep = " ") %>%
  separate(pausestart, into = c("baddate4", "pausestart"), sep = " ") %>%
  separate(pauseend, into = c("baddate5", "pauseend"), sep = " ") %>% 
  select(-contains("bad")) 
```
Combine date and time to form a dttm column and set the time zone to PHT, Asia/Manila
```{r}
surv$start <- str_c(surv$date, surv$starttime, sep = " ")  
surv$start <- ymd_hms(surv$start)
surv$start <- force_tz(surv$start, tzone = "Asia/Manila")
surv$end <- str_c(surv$date, surv$endtime, sep = " ")  
surv$end <- ymd_hms(surv$end)
surv$end <- force_tz(surv$end, tzone = "Asia/Manila")
surv$paust <- str_c(surv$date, surv$pausestart, sep = " ")  
surv$paust <- ymd_hms(surv$paust)
surv$paust <- force_tz(surv$paust, tzone = "Asia/Manila")
surv$pausend <- str_c(surv$date, surv$pauseend, sep = " ")  
surv$pausend <- ymd_hms(surv$pausend)
surv$pausend <- force_tz(surv$pausend, tzone = "Asia/Manila")
```
Change time zone to UTC 
```{r}
surv$start <- with_tz(surv$start, tzone = "UTC")
surv$end <- with_tz(surv$end, tzone = "UTC")
surv$paust <- with_tz(surv$paust, tzone = "UTC")
surv$pausend <- with_tz(surv$pausend, tzone = "UTC")
```
Read in each GPX file, find the survey that matches, and trim the file to fit the survey, write an output file in gpx format.  **TODO: This is taking a very long time to run.**  Need to find a better way to do this. **TODO: This is producing an NA file, figure out why**
```{r}
folders  <-  list.files(path = "./", pattern ="gps") # move any empty folders out of the data folder
for (l in 1:length(folders)){
  files <- list.files(path = paste("./",folders[l], sep = ""), pattern = "*Track*")
  for(i in 1:length(files)){ # for each file
	infile <- readGPXGarmin(paste("./", folders[l], "/", files[i], sep="")) 
	header <- infile$header
	data <- infile$data
	data$time <- ymd_hms(data$time)
	data <- arrange(data, time)
	# start time for this GPX track
	instarttime <-  data$time[1]
	# end time for this GPX track
	inendtime = data$time[nrow(data)]
	# change elevation to zero
	data$elev <- 0
	data$unit <- substr(folders[l],4,4)
	
	
	# which survey started after the gpx and ended before the gpx ####
	inds <- surv %>% 
	  filter(start >= instarttime & end <= inendtime & !is.na(divenum))
	
	# if none of the surveys fit ####
	if(nrow(inds) == 0){
		print(str_c("File", files[i], "does not cover a complete survey"))
	  
	  # find a survey that at least starts or ends within this GPX track
	  # which survey ended before the gpx and ended after the gpx started or started before the gpx ended and started after the gpx started
		inds <- surv %>% 
		  filter((end <= inendtime & end >= instarttime)| (start <= inendtime & start >= instarttime))
		if(nrow(inds) == 0){
			print(str_c("EVEN WORSE:", files[i], "does not cover even PART of a survey"))
		}
	}
	if(nrow(inds) > 0){
		for(j in 1:nrow(inds)){ # step through each survey that fits within this track (one or more)
			# output all: not just if this was a dive for collecting APCL
		  # if no pause
		  if(is.na(inds$paust[j])){ 
		    # find the GPX points that fit within the survey
				k <- data %>% 
				  filter(time >= inds$start[j] & time <= inds$end[j])
				k$lat <- as.character(k$lat)
				k$lon <- as.character(k$lon)
				k$elev <- as.character(k$elev)
				k$time <- as.character(k$time)
				outfile <- k
			
				
				writeGPX(filename = str_c("gpx_trimmed/GPS", outfile$unit[1], "_", inds$divenum[j], "_", files[i], sep=""), outfile = outfile)
			}
			if(!is.na(inds$paust[j])){ # account for a pause if need be
			  k1 <- data %>%
			    filter(time >= inds$start[j] & time <= inds$paust[j])
			  k1$lat <- as.character(k1$lat)
			  k1$lon <- as.character(k1$lon)
			  k1$elev <- as.character(k1$elev)
			  k1$time <- as.character(k1$time)
			  outfile1 <- k1

			  k2 <- data %>%
			    filter(time >= inds$pausend[j] & time <= inds$end[j])
			  k2$lat <- as.character(k2$lat)
			  k2$lon <- as.character(k2$lon)
			  k2$elev <- as.character(k2$elev)
			  k2$time <- as.character(k2$time)
			  outfile2 <- k2

			  writeGPX(filename = paste("gpx_trimmed/GPS", outfile$unit[1], "_",inds$divenum[j], "_", files[i], '_1.gpx', sep=''), outfile = outfile1) # write as two tracks
			  writeGPX(filename = paste("gpx_trimmed/GPS", outfile$unit[1], "_",inds$divenum[j], "_", files[i], '_2.gpx', sep=''), outfile = outfile2)
			  

			}
		}
	}
  }
}
```


