---
title: "Summarize fish for each site and dive"
author: "Michelle Stuart"
date: "6/7/2017"
output: github_document
---
### This is a script to determine for each dive the number of fish captured, sampled and recatpured and the same numbers at each site, also the number of dives per site.
```{r setup}
library(tidyverse)
library(stringr)
excel_file <- ("data/GPSSurveys2017.xlsx")
```

#### fish observation info from anem survey
```{r}
fish <- readxl::read_excel(excel_file, sheet = "anemones", col_names = TRUE, na = "")
names(fish) <- str_to_lower(names(fish))
fish <- filter(fish, !is.na(id)) %>% 
  distinct()
```

#### get dive info
```{r}
site <- readxl::read_excel(excel_file, sheet = "diveinfo", col_names=TRUE)
names(site) <- stringr::str_to_lower(names(site))
site <- site %>% filter(!is.na(divenum)) %>% select(divenum, site) %>% distinct()
```


#### fish processing info from clownfish survey
```{r}
samp <- readxl::read_excel(excel_file, sheet = "clownfish", col_names = TRUE, na = "")   
names(samp) <- stringr::str_to_lower(names(samp))
samp <- filter(samp, !is.na(divenum) & gps == 4)
```


#### make a table of observed fish
```{r}
divetot <- fish %>%
  group_by(divenum) %>% 
  filter(!is.na(numfish) & spp == "APCL") %>% 
  summarise(observed = sum(numfish))
```


#### add the site to the table
```{r}
divetot <- left_join(divetot, site, by = "divenum")
```


#### how many tissue samples were collected?  
```{r}
tissue <- samp %>% 
  filter(!is.na(finid)) %>% 
  group_by(divenum) %>% 
  summarize(fins=n())
divetot <- left_join(divetot, tissue, by = "divenum")
```


#### how many fish were captured
```{r}
cap <- samp %>% filter(!is.na(size)) %>% 
    group_by(divenum) %>% 
    summarise(captured = n())
divetot <- left_join(divetot, cap, by = "divenum")
```


#### how many fish were recaptures?
```{r}
recap <- samp %>% 
  filter(recap == "Y") %>% 
  group_by(divenum) %>% 
  summarise(recap = n())

divetot <- left_join(divetot, recap, by = "divenum") 
divetot <- divetot %>% 
  select(divenum, site, observed, captured, fins, recap) %>% 
  distinct()

```

#### are any captured more than observed?
```{r}
bad <- filter(divetot, captured > observed) # the one at visca is ok because it is ????

```


#### are there more fin clips than captured?
```{r}
bad <- filter(divetot, fins > captured)
```

#### are there ore recaps than captured?
```{r}
bad <- filter(divetot, recap > captured)
```


#### Fish per site -----------------------------------------------------------

#### fish observed by site
```{r}
fishobs <- left_join(fish, site, by = "divenum") %>% 
  distinct()
sitetot <- fishobs %>% 
  filter(spp == "APCL", !is.na(numfish)) %>% 
  group_by(site) %>% 
  summarise(observed = sum(numfish))
```


#### how many tissue samples were collected?  
```{r}
tissue <- left_join(samp, site, by = "divenum") %>% 
  distinct()
fins <- tissue %>% 
  filter(!is.na(finid)) %>% 
  group_by(site) %>% 
  summarise(finclip = n())

sitetot <- left_join(sitetot, fins, by = "site")
```


#### how many fish were captured
```{r}
cap <- tissue %>% 
  filter(!is.na(size)) %>% 
  group_by(site) %>% 
  summarize(captured = n())

sitetot <- left_join(sitetot, cap, by = "site")
```


#### how many fish were recaptures?
```{r}
recap <- tissue %>% 
  filter(recap == "Y") %>% 
  group_by(site) %>% 
  summarize(recap = n())

sitetot <- left_join(sitetot, recap, by = "site") 
sitetot <- sitetot %>% 
  select(site, observed, captured, finclip, recap)
```


#### are any captured more than observed?
```{r}
bad <- filter(sitetot, captured > observed) #### the one at visca is ok because it is ????
```

#### are there more fin clips than captured?
```{r}
bad <- filter(sitetot, finclip > captured)
```

#### are there ore recaps than captured?
```{r}
bad <- filter(sitetot, recap > captured)

View(sitetot)
View(divetot)

```

