---
title: "Ordering lab supplies"
author: "Michelle Stuart"
date: "6/13/2018, updated `r Sys.Date()`"
params:
  num_sample: 4416
  num_plate: 46
  num_pool: 92
output:
  html_document: default
  pdf_document: default
---
The number of plates is the number of samples divided by 96.  The number of pools is the number of plates times 2.

```{r setup, echo=FALSE, message=FALSE, include=FALSE}
library(tidyverse)
```


Calculate a budget for processing samples - there is a 2 plate minimum for the process

####Extract 
```{r, echo=FALSE}
(extr <- tibble(qiagen_kits = ceiling((params$num_plate)/12), 
              tips_300 = 7*(params$num_sample), 
              plates = (params$num_plate), 
              seals = (params$num_plate)) %>% 
   gather(item, quantity))
```

####Digest 
```{r, echo=FALSE}
(digest <- tibble(
  tips_100 = 2*params$num_sample, 
  PstI_ul = 0.2*(params$num_sample*1.1), 
    MluCI_ul = 2*(params$num_sample*1.1),
    seals = params$num_plate, 
  plates = params$num_plate
) %>% 
   gather(item, quantity))

```
####Clean plates
```{r, echo=FALSE}
(clean <- tibble(
  vol_samples = 50,
    ampure_ul = (params$num_sample*1.1)* 1.5 * vol_samples, 
  ampure_ml = ceiling(ampure_ul/1000),
    ethanol_ul = 400 * (params$num_sample * 1.1), 
  ethanol_ml = ceiling(ethanol_ul/1000),
    tips_300 = params$num_sample * 4,
    tips_100 = params$num_sample * 3,  
    seals = params$num_plate, 
  plates = params$num_plate) %>% 
   select(-vol_samples) %>% 
   gather(item, quantity))

```
####Quantification of digest plates

```{r, echo=FALSE, message=FALSE, warning=FALSE}
(pico <- tibble(
  final_vol_pico = 100*params$num_sample,
  grenier_plate = (params$num_plate + 1),
    vol_1x_TE = 0.995 * final_vol_pico,
    pico_uL = final_vol_pico - vol_1x_TE,
    tips_100 = 2*params$num_sample,
    tips_10 = params$num_sample,
    foil = params$num_plate, 
    seals = params$num_plate
  ) %>% 
  select(-final_vol_pico, -vol_1x_TE) %>% 
   gather(item, quantity))
```
# Prepare adapters
1uL pure stock goes into 25uL of working stock.  We receive about 250uL of pure stock (it is rehydrated depending on the concentration we recieve, 250 is the low end.)  So if we receive 250uL of pure stock, that is 6250uL of working stock

# Ligation
```{r, echo=FALSE}
(lig_prep <- tibble(
  plates = params$num_plate, 
  water_ul = 22.2*params$num_sample,
  robot_time_min = 15 * params$num_plate,
  robot_time_hr = robot_time_min/60,
  seals = 2*params$num_plate, # because we seal it at the robot and then seal it again at the reaction
  P2_ul = 2*params$num_sample, 
  ligase_ul = 0.8*params$num_sample, 
  P1_ul =  2*params$num_sample 
) %>% 
   gather(item, quantity))
```

# Pool the samples and 2xSPRI

```{r, echo=FALSE}
(first_spri <- tibble(
  vol_samples = 200,
  num_vials = 4*params$num_pool, 
    ampure_ul = (num_vials*1.1)* 1.5 * vol_samples, 
  ampure_ml = ceiling(ampure_ul/1000),
    ethanol_ul = 400 * (num_vials * 1.1), 
  ethanol_ml = ceiling(ethanol_ul/1000),
    tips_300 = num_vials * 4,
    tips_100 = num_vials * 3,  
    tubes = num_vials + params$num_pool) %>% 
   select(-vol_samples, -num_vials)%>% 
   gather(item, quantity))

```

```{r, echo=FALSE}
(second_spri <- tibble(
  vol_samples = 45*4,
    ampure_ul = (params$num_pool*1.1)* 1.5 * vol_samples, 
  ampure_ml = ceiling(ampure_ul/1000),
    ethanol_ul = 400 * (params$num_pool * 1.1), 
  ethanol_ml = ceiling(ethanol_ul/1000),
    tips_300 = params$num_pool * 4,
    tips_100 = params$num_pool * 3,  
    tubes = params$num_pool) %>% 
   select(-vol_samples)%>% 
   gather(item, quantity))

```
# Pippin
```{r}
(pippin <- tibble(
  cassette_num = ceiling(params$num_pool/5),
  tips_10 = params$num_pool,
  tips_100 = 3* params$num_pool,
  tubes = params$num_pool
) %>% 
   gather(item, quantity))
```

# Primer Prep
1uL of primer goes into each reaction.  5uL of pure stock goes into each working stock of 100uL.  If there are 300ul of pure stock, that makes 60 vials of 100uL each, for a total of 6000 reactions.


# PCR
```{r}
(pcr <- tibble(
  phusion_ul = 0.2*params$num_pool,
  f_primer = params$num_pool,
  r_primer = params$num_pool,
  plates = ceiling(params$num_pool*4/96),
  tips_10 = 2*params$num_pool*4,
  seals = plates
) %>% 
   gather(item, quantity))
```
# Qubit
```{r}
(qubit <- tibble(
  qubit_ul = (params$num_pool)+20
) %>% 
  gather(item, quantity))
```


# Baits
There are 4 pools per baits reaction 
16 reactions:   $3600
48 reactions:   $5990
96 reactions:   $9120

```{r}
(baits <- tibble(
  baits_rxns = params$num_pool/4
) %>% 
   gather(item, quantity))

```



####Merge these tables into one grocery list
```{r, echo=FALSE}
grocery_list <- rbind(baits, clean, digest, extr, first_spri, lig_prep, pcr, pico, pippin, second_spri, qubit) %>% 
  group_by(item) %>% 
  summarise(quantity = sum(quantity))
```


# Price list
```{r}
cost <- read_csv(
  "
  item, num_units, cost
  ampure_ml, 515, 7200
  baits_rxns, 48, 5990
  cassette_num, 10, 475
  ethanol_ml, 4000, 104
  f_primer, 6000, 20
  foil, 100, 50 
  grenier_plate, 40, 70
  ligase_ul, 250, 256
  MluCl_ul, 500, 233
  P1_ul, 6250,65.54
  P2_ul, 6250,54.73
  phusion_ul, 100,200
  pico_uL, 1000,410.97
  plates, 100,227
  PstI_ul,500,233
  qiagen_kits,1,3082
  qubit_ul,1250,97
  r_primer,6000,20
  robot_time_hr, 1, 100
  seals,100,78.80
  tips_10,960,126
  tips_100,960,121
  tips_300,960,121
  tubes,500,103
  water_ul,1000,45"
) %>% 
  mutate(cost_per_item = cost/num_units)
```

# add cost to grocery
```{r}
grocery <- grocery_list %>% 
  left_join(select(cost, item, cost_per_item), by = "item") %>% 
  filter(!is.na(cost_per_item)) %>% 
  mutate(total_cost = cost_per_item * quantity, 
         cost_per_sample = formatC(total_cost/params$num_sample, digits = 2))

supplies_total <- grocery %>% 
  summarise(total_cost_per_sample = sum(cost_per_sample, na.rm = T))
```


# Cost for sequencing
```{r eval=FALSE}
seq <- tibble(
  samp_per_lane = 576,
  
)
```

